\documentclass[dvipsnames]{article}
\usepackage{minted}
\input{../article_preamble.tex}
\title{IO Problem Set 1 (BLP)}
\author{Chris Ackerman\thanks{I worked on this problem set with Luna Shen, David Kerns, and Benedikt Graf}}
\date{\today}


\begin{document}
\maketitle
\section*{Problem 1}
\subsection*{Estimate the Model using OLS, with price and promotion as characteristics}

\begin{minted}{python}
  #1.1: Estimate using OLS with price and promotion as product characteristics.
res_log1 = smf.ols('Y ~ prices + prom_', data=otc_dataDf).fit()
\end{minted}
\begin{table}[htp]
  \centering
  \include{1-1.tex}
\end{table}
\newpage

\subsection*{Estimate the Model using OLS, with price and promotion as characteristics, and brand dummies}

\begin{minted}{python}
#1.2: Estimate using OLS with price and promotion as product characteristics and brand dummies.
res_log2 = smf.ols('Y ~ prices + prom_ + C(brand)', data=otc_dataDf).fit()
  \end{minted}
\begin{table}[htp]
  \centering
  \include{1-2.tex}
  \end{table}
  \newpage

\subsection*{Estimate the Model using OLS, with price and promotion as characteristics, and store-brand dummies}
\begin{minted}{python}
#1.3. Using OLS with price and promotion as product characteristics and store-brand 
#(the interaction of brand and store) dummies.

res_log3 = smf.ols('Y ~ prices + prom_ + C(brand)*C(store)', data=otc_dataDf).fit()
  \end{minted}
\begin{table}[htp]
  \centering
  \include{1-3.tex}
  \end{table}
  \newpage

\subsection*{Estimate the models from parts 1--3 using wholesale cost as an instrument}
\begin{minted}{python}
  # OLS with price and promotion as product characteristics, using wholesale cost as instrument
wholeSale_IV1 = iv.IV2SLS.from_formula('Y ~ 1 + [prices ~ cost_] + prom_ ', data=otc_dataDf).fit()
\end{minted}
\begin{table}[htp]
  \centering
  \include{1-4a.tex}
  \end{table}
 
  \newpage

  \begin{minted}{python}
    # OLS with price and promotion as product characteristics and brand dummies
# using wholesale cost as instrument

wholeSale_IV2 = iv.IV2SLS.from_formula('Y ~ 1 + [prices ~ cost_] + prom_ + C(brand)', data=otc_dataDf).fit()
\end{minted}
\begin{table}[htp]
  \centering
  \include{1-4b.tex}
  \end{table}

  \begin{minted}{python}
# OLS with price and promotion as product characteristics and brand dummies
# using wholesale cost as instrument

wholeSale_IV3 = iv.IV2SLS.from_formula('Y ~ 1 + [prices ~ cost_] + prom_ + C(brand)*C(store)', data=otc_dataDf).fit()
    \end{minted}
\begin{table}[htp]
  \centering
  \include{1-4c.tex}
  \end{table}

\newpage
  \subsection*{Estimate the models from parts 1--3 using the Hausman instrument}

\begin{table}[htp]
  \centering
  \include{1-5a.tex}
  \end{table}
  \newpage

\begin{table}[htp]
  \centering
  \include{1-5b.tex}
  \end{table}

\begin{table}[htp]
  \centering
  \include{1-5c.tex}
  \end{table}

\newpage
\newpage
\subsection*{Mean own-price elasticities from the estimates in models 1--3}

\begin{table}[htp]
  \centering
  \include{1-6.tex}
  \end{table}

These results make sense. As a rule of thumb, the own-price elasticities should be $\in (-2, -5)$. The IV estimates using the Hausman instrument are approximately in this range. The OLS estimates are not, indicating that endogeneity is a practical concern in this setting. The estimates with wholesale cost as an instrument are also ``too small'', indicating that wholesale cost may not be a viable instrument in this context.

\newpage
\section*{Problem 2}
Our results for this section don't make sense, and there are a few indications that we may have errors in our code. Our GMM function reached the maximum number of iterations. It's not clear if we set the tolerance level too low or if it simply failed to converge; if it didn't converge then the estimation procedure is incorrect and our results \emph{shouldn't} make sense. 
\subsection*{Estimate the parameter values using BLP}
\begin{table}[htp]
  \centering
  \begin{tabular}{ccccc}
    $\alpha$ & $\beta$ & $\sigma_{ib}$ & $\sigma_I$ & $\sigma_I^2$\\
    \hline
    $1.67671625$ & $\begin{bmatrix}
 40.5539681 \\
       -37.20820459\\
        22.10159763\\
        -9.87790967\\
       -18.56618279\\
        17.5039916 \\
       -32.63598696\\
        34.20049778\\
       -12.49605257\\
         0.95162148\\
       -53.64393846\\
    \end{bmatrix}$
             & $\begin{bmatrix}
0.35372042 \\ 0.4527695 \\ 0.2383683
               \end{bmatrix}$ & $0.005161$ & $-0.121263$
  \end{tabular}
\end{table}
Note that we have one more element of $\beta$ than we should. When we constructed the dummies, we neglected to exclude one dummy variable, but we were unable to re-run our code before submitting our results. Given more time, we would re-run this step, and we suspect that the coefficients would then make more sense.

\subsection*{What are the elasticities for store 9 in week 10?}
We under-estimated the difficulty of this question. We know that we can calculate this from the first order conditions, and that we need to take a numerical integral over consumers, but we don't know how to do this.

\subsection*{Back out the marginal costs for store 9 in week 10. How are they different from wholesale costs?}

We have code that does this (for part 3) but we were unable to run it before turning in the HW. It should be almost exactly the same procedure as in part 3, but we need to use the random coefficient model's coefficients, and we are running into bugs in the code when we do that.

\newpage
\section*{Problem 3}
\subsection*{Predict the post-merger prices using the logit model, but only for store 9 in week 10}

\begin{table}[htp]
  \centering
  \include{3-1.tex}
\end{table}

\subsection*{How to predict the change in prices after the merger using the random coefficients model?}

The procedure for predicting the effects of a merger using estimates from the random coefficients model is exactly the same. When we estimate the logit model, we end up with unrealistic elasticities, and hence unrealistic substitution patterns. The random coefficients model gives us a more reasonable matrix of elasticities, but we use it in the same way. When we want to assess the change of moving from two single-product firms to a multi-product firm, we switch from looking at firms that take FOCs with respect to a single price to a single firm that takes FOCs with respect to two prices (the price of each good that it produces). What does this mean? With a more realistic estimate of substitution patterns, we have a better idea of how consumers will respond to changes in prices. Pre-merger, a firm would be ``unwilling'' to raise its price because it would lose customers. Now, however, the merged firm \emph{may} recognize that increasing the price of one of it's products will cause consumers to switch to its other product, so its profit maximizing price may be higher in the merged case than in the pre-merger case. The random coefficients model is giving us a better insight into these actual substitution patterns.
\end{document}
%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
